H52F-05H: Robust Adaptation to\\Multi-Scale Climate Variability

### 2018-12-10 slides-agu/presentation.tex
Update slides -- new look

\documentclass[
  10pt,     % set default generic font size
  %handout   % ignores all the \pause commands
]{beamer}
\usetheme[block=fill]{metropolis}

@@

@@

\section{Setup}

\begin{frame}{NYC After Sandy \citep{CityofNewYork:2013uh}}
  \begin{columns}[T]
    \pause
    \begin{column}{0.5\textwidth}
      \begin{figure}
        \includegraphics[width=\columnwidth]{three-barriers.png}
      \end{figure}
      \begin{itemize}
        \item Environmental, social impacts
        \item Permanent: $\approx \$25 \text{billion}$ ``locked-in''
      \end{itemize}
    \end{column}
    \pause
    \begin{column}{0.5\textwidth}
      \begin{figure}
        \includegraphics[width=\columnwidth]{flood-concept.png}
      \end{figure}
      \begin{itemize}
        \item Targeted local resilience $\Uparrow$ and vulnerability $\Downarrow$
        \item Portfolio of investments
        \item Update regularly
      \end{itemize}
    \end{column}
  \end{columns}
  \pause
  \begin{alertblock}{How should NYC allocate resources, over time,}
    for infrastructure and climate adaptation given uncertainty in future climate and societal needs?
  \end{alertblock}
\end{frame}

\begin{frame}{Idea 1: Risk Estimates over Finite Future Periods}
  \begin{alertblock}{Typical Approach:}
    \gls{cba} over a \alert{finite} planning horizon of $M$ years (explicitly or implicitly).
  \end{alertblock}
  \begin{columns}
    \pause
    \begin{column}{0.55\textwidth}
      Thus, evaluation of project depends on climate conditions over the next $M$ years:
      \begin{itemize}
        \item For ``mega-project'', $M \geq \SI{50}{years}$
        \item For small, flexible project, $M \leq \SI{5}{years}$
      \end{itemize}
      \pause
      So, misrepresentation of climate risk over $M$ years $\rightarrow$ bias in \gls{cba}
    \end{column}
    \pause
    \begin{column}{0.4\textwidth}
      \includegraphics[width=\textwidth]{Bias-Variance-Sketch}
      {\small Estimation bias and variance drive project success!}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Idea 2: Physical Drivers of Risk Depend on $M$}
  The physical drivers of hazard depend on the projection horizon ($M$).
  \begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio=true]{conceptual-sketch-a.png}\\
  \end{figure}
  \pause
  However, our ability to identify these drivers depends on information available (\eg, the length of an $N$-year observational record).
  \begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio=true]{conceptual-sketch-b.png}
  \end{figure}
  \emph{(informational uncertainty $\Leftarrow$ clouds, sea ice, subsurface, \etc)}
\end{frame}

\begin{frame}{Research Objective}
  \begin{alertblock}{How well can one identify and predict risk}
     associated with cyclical and secular climate signals over a finite design life ($M$),  given limited information (\eg, an $N$-year record).
  \end{alertblock}
  \pause
  In particular, we consider:
  \begin{itemize}
    \item the specific roles of $M$ and $N$;
    \pause\item the choice of estimation strategy; and
    \pause\item the form of the underlying climate signal.
  \end{itemize}
  \pause
  Use simplest possible representation of key processes.
\end{frame}

@@

\begin{frame}{Breathless Overview of Methods}
  \begin{columns}[T]
    \begin{column}{0.475\textwidth}
      We will consider:
      \begin{itemize}
        \item 3 scenarios:
        \begin{itemize}
          \item \gls{lfv} only
          \item linear trend only
          \item trend + \gls{lfv}
        \end{itemize}
        \item 3 fitting models:
        \begin{itemize}
          \item stationary
          \item trend
          \item \gls{hmm}
        \end{itemize}
        \item Many values of $\qty{M, N}$
      \end{itemize}
    \end{column}
    \pause
    \begin{column}{0.475\textwidth}
      \begin{itemize}
        \item \gls{lfv} from \gls{enso} model \citep{Ramesh:2016hf, Zebiak:1987cl}
        \item For each combination, use \num{1000} synthetic streamflow sequences to evaluate the expected bias and variance (recall: these map to bias of \gls{cba})
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Stationary Scenario (LFV Only)}
  Estimating a trend when none exists may lead to substantial over-estimation of risk, particularly for short $N$ or long $M$.
  Explicitly modeling \gls{lfv} improves estimates, particularly for long $N$.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Nonstationary Scenario I (Secular Change Only)}
  For long $M$, it's important to estimate the trend -- but only works well if $N$ also long.
  For short $M$ and short $N$, the uncertainty added by estimating a trend is not worth the bias reduction.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Nonstationary Scenario II (Secular Change + LFV)}
  Conclusions are similar to ``Secular Change Only'' but variance has increased -- \ie as the system becomes more complex, more data is needed to understand it.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.7\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
  \end{figure}
\end{frame}

@@

\begin{frame}{Key Findings}
  \begin{alertblock}{How should NYC allocate resources, over time,}
    for infrastructure and climate adaptation given uncertainty in future climate and societal needs?
  \end{alertblock}
  \begin{itemize}
    \pause
    \item When planning period is short, \gls{lfv} dominates \citep{Hodgkins:2017hw} $\Rightarrow$ uncertainty small
    \pause
    \item When $M$ long, future changes must be estimated using complex models which require more information $\Rightarrow$ more uncertainty
    \pause
    \item As uncertainty increases, adaptation strategies with short planning periods are preferred
  \end{itemize}
\end{frame}

\begin{frame}{A Shameless Pitch!}
  I'm working on turning these concepts for project \emph{evaluation} into models for \emph{decision-making}.
  Specifically:
  \pause
  \begin{itemize}
    \item Integrating paleo records, observations, and \glspl{gcm} for climate risk projection
    \pause
    \item Valuing the ``option'' to build a particular project
    \pause
    \item Optimizing portfolios of long- and short-term investments
  \end{itemize}
  \pause
  If you have a working model of your system and are making long-/short-term investment choices I'd \alert{love to talk} further!
\end{frame}

@@

>>>>

\documentclass[
  10pt,     % set default generic font size
  handout   % ignores all the \pause commands
]{beamer}
\usetheme[block=fill]{metropolis}

@@

% Add footer for AGU
\setbeamertemplate{footline}[text line]{%
  \parbox{0.8\linewidth}{
    \vspace*{-8pt}James Doss-Gollin (\url{james.doss-gollin@columbia.edu})
  }
  \hfill%
  \parbox{0.15\linewidth}{
    \vspace*{-8pt}\raggedleft\insertpagenumber
  }
}

@@

\begin{frame}{Motivating Example}
  \begin{columns}[T]
    \pause
    \begin{column}{0.5\textwidth}
      \begin{figure}
        \centering
        \caption{
          Permanent structural solution would pose environmental and social impacts and lock in $\approx \$25 \text{billion}$;
          a portfolio of smaller projects may allow a more flexible strategy.
        }
        \includegraphics[width=0.85\columnwidth]{three-barriers.png}\\
        \includegraphics[width=0.85\columnwidth]{flood-concept.png}
      \end{figure}
    \end{column}
    \pause
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item ``Because present decisions impact a system's ability to adapt to future needs, flexibility in activating, delaying, and replacing engineering projects should be considered in least‐cost water supply intervention scheduling,'' \citep{erfani:2018}.
        \item How to build a portfolio of risk management investments, over time, given uncertain future climate/society/technology?
      \end{itemize}
    \end{column}
  \end{columns}
\end{frame}

\section{Underlying Ideas}

\begin{frame}{Idea 1: Risk Estimates over Finite Future Periods}
  \begin{alertblock}{Typical Approach:}
    \gls{cba}, probably with discounting, over a \alert{finite} planning horizon of $M$ years (explicitly or implicitly).
  \end{alertblock}
  \begin{columns}
    \pause
    \begin{column}{0.5\textwidth}
      Thus, evaluation of project depends on climate conditions over the next $M$ years:
      \begin{itemize}
        \item For ``mega-project'', $M \geq \SI{50}{years}$
        \item For small, flexible project, $M \leq \SI{5}{years}$
      \end{itemize}
    \end{column}
    \pause
    \begin{column}{0.5\textwidth}
      Misrepresentation of climate risk over $M$ years $\rightarrow$ bias in \gls{cba}
      \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio=true]{Bias-Variance-Sketch}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Idea 2: Physical Drivers of Risk Depend on $M$}
  The physical drivers of hazard depend on the projection horizon ($M$).
  \begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio=true]{conceptual-sketch-a.png}\\
  \end{figure}
  \pause
  However, our ability to identify these mechanisms depends on information available (\eg, the length of an $N$-year observational record).
  \begin{figure}
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio=true]{conceptual-sketch-b.png}
  \end{figure}
\end{frame}

@@

\begin{frame}{Research Objective}
  \begin{alertblock}{How well can one identify and predict risk}
     associated with cyclical and secular climate signals over a finite design life ($M$),  given limited information (\eg, an $N$-year record).
  \end{alertblock}
  \pause
  In particular, we consider:
  \begin{itemize}
    \item the specific roles of $M$ and $N$;
    \pause
    \item estimation strategy (few processes/parameters $\Rightarrow$ many processes/parameters)
    \pause
    \item the form of the underlying climate signal.
  \end{itemize}
  \pause
  Use simplest possible representation of key processes.
\end{frame}

\begin{frame}{Breathless Overview of Methods}
  \begin{columns}[T]
    \begin{column}{0.475\textwidth}
      We will consider:
      \begin{itemize}
        \item 3 scenarios:
        \begin{itemize}
          \item \gls{lfv} only
          \item linear trend only
          \item trend + \gls{lfv}
        \end{itemize}
        \item 3 fitting models:
        \begin{itemize}
          \item stationary
          \item trend
          \item \gls{hmm}
        \end{itemize}
        \item Many values of $\qty{M, N}$
      \end{itemize}
    \end{column}
    \pause
    \begin{column}{0.475\textwidth}
      \begin{figure}
        \caption{Over- or under-estimating uncertainty leads to biased estimates of $p(X > X^*)$ even if estimate of $X$ is unbiased}
        \includegraphics[width=\textwidth]{Example-NINO3-M100-N50-LN2.jpg}
      \end{figure}
      For each combination, use \num{1000} synthetic streamflow sequences to evaluate the expected bias and variance of $p(X > X^*)$
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{Stationary Scenario (LFV Only)}
  Estimating a trend when none exists may lead to substantial over-estimation of risk, particularly for short $N$ or long $M$.
  Explicitly modeling \gls{lfv} improves estimates, particularly for long $N$.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.625\textheight,keepaspectratio=true]{lfv-only-nino3-bias-variance.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Nonstationary Scenario I (Secular Change Only)}
  For long $M$, it's important to estimate the trend -- but only works well if $N$ also long.
  For short $M$ and short $N$, the uncertainty added by estimating a trend is not worth the bias reduction.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.625\textheight,keepaspectratio=true]{secular-only-nino3-bias-variance.pdf}
  \end{figure}
\end{frame}

\begin{frame}{Nonstationary Scenario II (Secular Change + LFV)}
  Conclusions are similar to ``Secular Change Only'' but variance has increased -- \ie as the system becomes more complex, more data is needed to understand it.
  \begin{figure}
    \centering
    \includegraphics[width=\textwidth,height=0.625\textheight,keepaspectratio=true]{lfv-secular-nino3-bias-variance.pdf}
  \end{figure}
\end{frame}

@@

\begin{frame}{Summary}
  Starting point:
  \begin{enumerate}
    \item Investment evaluation depends on climate condition over finite planning period
    \item Physical drivers of risk depend on planning period
  \end{enumerate}
  \pause
  As a consequence:
  \begin{itemize}
    \item \emph{Ability} to identify and predict different climate signals depends on information available (\eg, $N$)
    \item \emph{Importance} of identifying and predicting different climate signals depends on degree of extrapolation desired (\ie, $M$)
  \end{itemize}
  \pause
  In general, low risk tolerance and/or short $N$ favor short-$M$ investments relative to long-$M$ ones.
\end{frame}

\begin{frame}{A Shameless Pitch!}
  I'm working on turning these concepts for project \emph{evaluation} into models for \emph{decision-making}.
  Specifically:
  \begin{itemize}
    \item Integrating paleo records, observations, and \glspl{gcm} for climate risk projection
    \item Valuing the ``option'' to build a particular project
    \item Optimizing portfolios of long- and short-term investments
  \end{itemize}
  If you have a working model of your system and are making long/short-term investment choices I'd \alert{love to talk} further!
\end{frame}

@@

\begin{frame}{Example Sequences and Fits}
  \begin{figure}
    \includegraphics[width=\textwidth]{Example-NINO3-M100-N50.pdf}
    \caption{Example of sequences generated with $M=100$ and $N=50$}
  \end{figure}
\end{frame}
